---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data

```{r, echo=TRUE}
setwd("C:/Users/caozhiy/Documents/data/ds5/project1")
originalData <- read.csv("activity.csv")
originalData$date <- as.Date(originalData$date)
```
## What is mean total number of steps taken per day?
```{r, echo=TRUE}
library(reshape2)
summaryByDateMelt <- melt(originalData, id.vars="date", measure.vars="steps", na.rm=FALSE)
summaryByDateCast <- dcast(summaryByDateMelt, date ~ variable, sum)
plot(summaryByDateCast$date, summaryByDateCast$steps, type="h", main="Daily Steps", xlab="Date", ylab="Steps per Day", col="red", lwd=8)
mean(summaryByDateCast$steps, na.rm=TRUE)
median(summaryByDateCast$steps, na.rm=TRUE)
```


## What is the average daily activity pattern?
```{r, echo=TRUE}
summaryByIntervalMelt <- melt(originalData, id.vars="interval", measure.vars="steps", na.rm=TRUE)
summaryByIntervalCast <- dcast(summaryByIntervalMelt, interval ~ variable, mean)
plot(summaryByIntervalCast$interval, summaryByIntervalCast$steps, type="l", main="Average Steps in each Interval", xlab="Interval", ylab="Steps", col="blue", lwd=2)

summaryByIntervalCast$interval[which(summaryByIntervalCast$steps == max(summaryByIntervalCast$steps))]
```

## Imputing missing values
```{r, echo=TRUE}
joinedData <- merge(originalData, summaryByIntervalCast, by="interval")
naIndex <- which(is.na(joinedData$steps.x))

joinedData[naIndex, "steps.x"] <- joinedData[naIndex, "steps.y"]

imPutedData <- joinedData[c("interval", "steps.x", "date")]
names(imPutedData) <- c("interval", "steps", "date")

impuSumByDateMelt <- melt(imPutedData, id.vars="date", measure.vars="steps", na.rm=FALSE)
impuSumByDateCast <- dcast(impuSumByDateMelt, date ~ variable, sum)

plot(impuSumByDateCast$date, impuSumByDateCast$steps, type="h", main="Daily Steps with Imputed Data", xlab="Date", ylab="Steps", lwd=5)
mean(impuSumByDateCast$steps)
median(impuSumByDateCast$steps, na.rm=TRUE)
```
Looks like the both mean and median doesn't change much from previous calculated without imputing.

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
imPutedData$weekday <- weekdays(imPutedData$date)
imPutedData$dayType <- ifelse(imPutedData$weekday %in% c("Saturday", "Sunday"), "Weekend", "Weekday")

subData <- imPutedData[c("interval", "dayType", "steps")]

dayTypeSummaryMelt <- melt(subData, id=c("interval", "dayType"), na.rm=TRUE, measures.vars="steps")
dayTypeSummaryCast <- dcast(dayTypeSummaryMelt, interval + dayType ~ variable, mean)

library(ggplot2)
p <- ggplot(dayTypeSummaryCast, aes(x=interval, y=steps)) + geom_line() + ylab("Number of Steps")
p <- p + facet_wrap(~ dayType, ncol=1)
p
```


Looks like the weekend will have average high numbers of steps.
